---
title: "Appetite Analysis - Phenotypic Plasticity at the Initiation of Dominance hierarchies"
date: "`r Sys.Date()`"
output:
  rmdformats::downcute:
    self_contained: true
    default_style: "light"
    downcute_theme: "default"
    code_folding: hide
    toc_float: FALSE
---

<style>
  h2{color: #DC7633 !important}
  h1{color: #5499C7 !important}
  body{background-color: white !important}
</style>

<style>
.list-group-item.active, .list-group-item.active:hover, .list-group-item.active:focus {
    color: #212F3D;
    background-color: #EDBB99;
    font-weight: bold;}
a {
    color: #5499C7;}
.nav-tabs > li.active > a, .nav-tabs > li.active > a:hover, .nav-tabs > li.active > a:focus {
    color: #DC7633;}
body {
    font-family: "Open Sans","Helvetica Neue",Helvetica,Arial,sans-serif;
    font-size: 15px;
    line-height: 1.42857143;
    color: #212F3D;}
</style>



<br/>

```{r set some defaults, echo = FALSE}

library("knitr")
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
pdf.options(useDingbats = FALSE)
options(knitr.kable.NA = '')

```

```{r load packaged, echo=FALSE}
library(ggplot2)
library(plyr)
library(dplyr)
library(multcomp)
library(tidyr)
library(lmerTest)
library(emmeans)
library(interactions)
library(kableExtra)
library(MuMIn)
```

## Introduction
This Rmarkdown document contains analysis of appetite of fish *Amphiprion percula* which were part of a 5 week experiment. Individuals were size-matched and either housed as a pair or as a singleton. A total of 30 replicates containing 30 pairs (n=60) and 30 singletons (n=30) were used in this experiment.

All data were obtained from video recordings of 5 minutes length. Pairs were separated with a divider, prior to feeding to avoid any unequal food intake due to bullying. Dividers were placed in singleton tanks as a sham control. From first bite taken, total number of bites over 5 minutes were recorded weekly for all individuals. There were no videos made of singletons during week 1. It was not possible to tell pairs apart, therefore, within pairs appetite was recorded for left and right fish. 

LmerTest package was used for all general linear mixed model analysis (GLMM).

---

## Inspecting the dataset {.tabset}
```{r, include=FALSE}
beh_data <- read.csv(file = "../Data/Strategic_Growth_Behavioural_DATA_20230202.csv", header = T)
head(beh_data)
str(beh_data)
dim(beh_data)
#many columns - must extract columns of interest
#extracting columns of interest:
#replicate_ID; housing_tank_ID; experiment_day; week; fish_type; abs_diff_appetite; average_appetite; total_appetite; P_max_appetite; P_min_appetite; S_appetite

beh_data <- beh_data[, c("replicate_ID", "housing_tank_ID", "experiment_day", "exp_week", "fish_type",  "abs_diff_appetite", "average_appetite", "total_appetite", "P_max_appetite", "P_min_appetite", "S_appetite")]
head(beh_data)

beh_data$abs_diff_appetite <- as.numeric(beh_data$abs_diff_appetite)
beh_data$average_appetite <- as.numeric(beh_data$average_appetite)
head(beh_data) #have columns of interest

#needing to remove 3 replicates which lost fish during the experiment
beh_data <- subset(beh_data, !replicate_ID == "8" & !replicate_ID == "9" & !replicate_ID == "18")
```

### List of replicates to use
```{r}
dim(beh_data)
print(beh_data$replicate_ID)
print(table(beh_data$fish_type))
```
<br/>

Appetite can be analysed in two different manners: <br/>
- Having pairs combined: Comparison of per capita appetite between pairs vs singleton <br/>
- Having pairs separated out: Comparison across three individuals <br/>
  - P-max (the individual within the pair that ate more)
  - P-min (the individual within the pair that ate less)
  - S (singleton)

Hypothesis: We would expect to see pairs eat more compared to singletons as they attempt to out compete each other. <br/>
Within pairs, we would expect 1 individual to eat more compared to its tank mate. 
Analysing across 3 fish types

---

## Pairs separated out {.tabset}
Previously we had a per capita appetite for both pairs combined, however, this is not the most informative when wanting to understand how is appetite driven within pairs. **Does one individual within the pair account for all that increase we see compared to singleton?**

Given we cannot tell individuals apart in the pairs, we must assign P-max and P-min, to any given individual within the pair at any given time. Whoever took more bites within the pair will be denoted as P-max and whoever ate less denoted as P-min for any given observation. We can only assume that whoever ate more at week one would be eating more throughout the experiment, however, there is no way to double-check this. But doing this will allow us to compare appetite across all fish categories, making a distinction amongst pairs.

needing to create a dataset for number of bites: 
- P_max & P_min in pairs and S for singleton
- needing columns: rep ID, housing_tank_ID, exp_week, fish_type, P_max_appetite, P_min_appetite, S_appetite


```{r preparing dataset, include=FALSE}
head(beh_data)

beh_data_app <- beh_data[, c("replicate_ID", "housing_tank_ID", "exp_week", "fish_type", "P_max_appetite", "P_min_appetite", "S_appetite")]
head(beh_data_app)
dim(beh_data_app)    #235   rows, I`ll need three times of this

beh_data_app <- gather(beh_data_app, key =  "appetite_type", value = "number_of_bites", 5:7)
head(beh_data_app)
dim(beh_data_app)   #705  rows
```
### Mean number of bites of P-max, P-min, S 
```{r, include=FALSE}
# ddply provides the standard deviation, standard error of the mean, and a (default 95%) confidence interval
beh_app_summary <- ddply(beh_data_app, c("exp_week", "appetite_type"),summarise,
               N    = sum(!is.na(number_of_bites)),
               mean = mean(number_of_bites, na.rm=TRUE),
               sd   = sd(number_of_bites, na.rm=TRUE),
               se   = sd / sqrt(N) )
beh_app_summary
```

```{r}
#there is a NaN again, needs fixing
beh_app_summary$N[beh_app_summary$N == 0] <- NA
beh_app_summary$mean[beh_app_summary$mean == "NaN"] <- NA

knitr::kable(beh_app_summary, caption = "Summary Statistics of Appetite", align = "c")

#appetite_type as factor:
beh_data_app$appetite_type <- as.factor(beh_data_app$appetite_type)
```

```{r}
ggplot(beh_app_summary%>% filter(!is.na(mean)), aes(x=exp_week ,y=mean, colour = appetite_type)) + geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.1) + geom_line(aes(group=appetite_type)) + geom_point() +   theme(legend.position="right") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black")) 
```

This figure suggest that both individual within the pair eats more at the beginning of the experiment compared to the singletons. Let`s look at linear-mixed models.

### Linear Mixed Model II
$formula: NumberOfBites$ ~ $AppetiteType*ExpWeek + (1|replicateID)$
```{r}
mixed_mod_minmax_app <- lmer(number_of_bites~exp_week*appetite_type + (1|replicate_ID), data = beh_data_app)
anova(mixed_mod_minmax_app)
```

*RESULTS*: We can see that time and fish category and the interaction between them are significant predictors of appetite of an individual. Given the interaction between the 2 explanatory variables are significant, we should not report main effects. 


### Post-Hoc Contrasts 
```{r}
emm_mix_mod_minmax_app <- emmeans(mixed_mod_minmax_app, ~exp_week*appetite_type)
pairs(emm_mix_mod_minmax_app, simple = "each")

emms_minmax_tbl <- pairs(emm_mix_mod_minmax_app, simple = "each") %>%
  rbind() 

emms_minmax_tbl%>%
   kbl(col.names = c("Fish Category", "Time", "Contrast", "Estimate", "SE", "df", "t ratio", "p-value"), caption = "Simultaneous pair-wise post-hoc comaprisons for appetite of fish categories (P-max, P-min, S) and time (week 1 to 5).", align = "c", digits = 3) %>%
  kable_classic(full_width = F, html_font = "Cambria")%>%
  column_spec(1:3, bold = T) %>%
  row_spec(0, color = "white", background = "black",bold = T)
```
>For manuscript: GLMM results showed that the interaction between fish type (P-max vs P-min vs S) and experimental week were significant predictors of appetite (F-value=2.2745, NumDf=7, DenDf=275.01, p-value=0.02885). Simultaneous pairwise post-hoc comparisons using Tukey’s HSD test showed significant difference between P-max and P-min as well as P-max and S appetite thoughout all 5 weeks (Table x). On the other hand, significant difference in appetite between P-min and S was significantly different only until week 3, and showed no statistical difference on week 4 and 5 (Table x). Within P-max, post-hoc pairwise comparison was significantly difference when comparing week 1 to weeks 2, 4, and 5 and when comparing week 3 to weeks 4 and 5 (Table x). Same results were found for P-min as well, while S showed no significant difference in appetite across the 5 weeks.

R-squared values of the best fit model:
```{r}
r.squaredGLMM(mixed_mod_minmax_app)
```

```{r}
cat_plot(mixed_mod_minmax_app, pred = exp_week, modx=appetite_type, geom = "line", interval=TRUE)
```


```{r}
mixed_mod_average_app_s <- step(mixed_mod_average_app,reduce.fixed = FALSE, reduce.random = FALSE)
mixed_mod_average_app_s
```

### Figures
```{r, fig.height=7, fig.width=7}
head(beh_app_summary)
colnames(beh_app_summary)[4] = "number_of_bites"

gg_min_max_appetite <- ggplot(beh_data_app, aes(x=exp_week, y=number_of_bites, group=appetite_type, colour=appetite_type))+
        geom_jitter(position = position_jitterdodge(0.2), size=4, alpha=0.25) +
        geom_line(aes(group=appetite_type), data = beh_app_summary,  position=position_dodge(0.7), size=1)+
        geom_pointrange(aes(ymin = number_of_bites-se, ymax = number_of_bites+se), data =beh_app_summary, size=1.2, alpha=1, position=position_dodge(0.7))+
        scale_color_manual(values = c( "magenta3","purple3", "grey50"), name ="social position", labels=c("P max", "P min", "S")) +
        theme_classic() +
        theme(axis.text = element_text(size = 18), axis.title = element_text(size = 18), legend.position = c(0.85,0.9),         legend.title=element_text(size=16),  legend.text=element_text(size=16))+
        ylab("Appetite (number of bites) + SEM") +
        xlab("Experimental week")

gg_min_max_appetite

#ggsave("../Figures/min_max_appetite-20250428-social-rank.png", height = 7, width = 7, dpi = 1200)
```
