---
title: "Aggression/Submission - Phenotypic Plasticity at the Initiation of Dominance hierarchies"
date: "`r Sys.Date()`"
output:
  rmdformats::downcute:
    self_contained: true
    default_style: "light"
    downcute_theme: "default"
    code_folding: hide
---
<style>
  h2{color: #DC7633 !important}
  h1{color: #5499C7 !important}
  body{background-color: white !important}
</style>

<style>
.list-group-item.active, .list-group-item.active:hover, .list-group-item.active:focus {
    color: #212F3D;
    background-color: #EDBB99;
    font-weight: bold;}
a {
    color: #5499C7;}
.nav-tabs > li.active > a, .nav-tabs > li.active > a:hover, .nav-tabs > li.active > a:focus {
    color: #DC7633;}
body {
    font-family: "Open Sans","Helvetica Neue",Helvetica,Arial,sans-serif;
    font-size: 15px;
    line-height: 1.42857143;
    color: #212F3D;}
</style>


<br/>

```{r set some defaults, echo = FALSE}

library("knitr")
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
pdf.options(useDingbats = FALSE)
options(knitr.kable.NA = '')

```

```{r load packaged, echo=FALSE}
library(ggplot2)
library(plyr)
library(dplyr)
library(multcomp)
library(tidyr)
library(lmerTest)
library(emmeans)
library(interactions)
library(MuMIn)
```

## Introduction
This Rmarkdown document contains analysis of behaviour (aggression & submission) of fish *Amphiprion percula* which were part of a 5 week experiment. Individuals were size-matched and either housed as a pair or as a singleton. A total of 30 replicates containing 30 pairs (n=60) and 30 singletons (n=30) were used in this experiment. This below analysis only concerns within pairs behaviour change over the 5 weeks.

All data were obtained from video recordings of 5 minutes length. For most pairs we could not tell them apart, therefore, aggression and submission cannot be directly correlated with dominant and subordinate fish. During the first 3 weeks, observed aggression and submission was in most instances displayed by both individual. From weeks 3-5, size difference has emerged between pairs and aggression was mainly displayed by bigger individual and submission was mainly a response to that. 

LmerTest package was used for all general linear mixed model analysis (GLMM).

---

## Inspecting the dataset {.tabset}
```{r, include=FALSE}
beh_data <- read.csv(file = "../Data/Strategic_Growth_Behavioural_DATA_20230202.csv", header = T)
head(beh_data)
str(beh_data)
dim(beh_data)
#many columns - must extract columns of interest
#extracting columns of interest:
#"replicate_ID", "housing_tank_ID", "experiment_day", "exp_week", "fish_type", "total_aggresion", "total_submission"

beh_pairs <- beh_data[, c("replicate_ID", "housing_tank_ID", "experiment_day", "exp_week", "fish_type", "total_aggresion", "total_submission")]
head(beh_pairs)


head(beh_pairs) #have columns of interest

#needing to remove 3 replicates which lost fish during the experiment
beh_pairs <- subset(beh_pairs, !replicate_ID == "8" & !replicate_ID == "9" & !replicate_ID == "18")

#this dataset has NA inputted for submission and aggression for singletons, we need to remove these so we`re only left with pairs
table(beh_pairs$fish_type)

beh_pairs <- subset(beh_pairs, fish_type == "P")
table(beh_pairs$fish_type) #great now we only have the pairs
```
### List of replicates to use
```{r}
dim(beh_pairs)
print(beh_pairs$replicate_ID)
print(table(beh_pairs$fish_type))
```
<br/>

Behaviour can be analysed as: <br/>
- aggression over time <br/>
- submission over time <br/>

Hypothesis: Over-time aggression and submission will decline as size difference emerges within pairs.

### Summary Statistics
```{r}
# ddply provides the standard deviation, standard error of the mean, and a (default 95%) confidence interval
agg_summary <- ddply(beh_pairs, c("exp_week"),summarise,
               N    = sum(!is.na(total_aggresion)),
               mean = mean(total_aggresion, na.rm=TRUE),
               sd   = sd(total_aggresion, na.rm=TRUE),
               se   = sd / sqrt(N) )

knitr::kable(agg_summary, caption = "Summary Statistics of Total Aggression", align = "c")

sub_summary <- ddply(beh_pairs, c("exp_week"),summarise,
               N    = sum(!is.na(total_submission)),
               mean = mean(total_submission, na.rm=TRUE),
               sd   = sd(total_submission, na.rm=TRUE),
               se   = sd / sqrt(N) )

knitr::kable(sub_summary, caption = "Summary Statistics of Total Submission", align = "c")
```

```{r, include=FALSE}
beh_pairs_totaggsub <- gather(beh_pairs, key = "behaviour_type", value = "count", 6:7 )
head(beh_pairs_totaggsub)

#getting summary results of stacked data:
pair_summary_totaggsub <- ddply(beh_pairs_totaggsub, c("behaviour_type", "exp_week"),summarise,
               N    = sum(!is.na(count)),
               mean = mean(count, na.rm=TRUE),
               sd   = sd(count, na.rm=TRUE),
               se   = sd / sqrt(N) )
pair_summary_totaggsub
```

```{r}
ggplot(pair_summary_totaggsub%>% filter(!is.na(mean)), aes(x=exp_week ,y=mean, colour = behaviour_type)) + geom_errorbar(aes(ymin=mean-se, ymax=mean+se), width=.1) +
    geom_line(aes(group=behaviour_type)) + geom_point(size=3) +   theme(legend.position="right") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.line = element_line(colour = "black"),axis.text = element_text(size = 25), axis.title = element_text(size = 20), legend.position = c(0.8,0.9), legend.title=element_text(size=18),  legend.text=element_text(size=18), legend.key = element_rect("white")) + xlab("experimental week") + ylab("total aggression/submission mean") + scale_color_manual(values = c("purple3", "plum3"), name="behaviour type", labels=c("aggression", "submission")) + ylim(0,10) 
```

Looking at this, both aggression and submission seem to decline over time. Let`s test this with a LMM.

---

## Aggression {.tabset}
$formula: TotalAggression$ ~ $time + (1|HousingTankID)$
```{r, include=FALSE}
str(beh_pairs)

#exp_week to factor:
beh_pairs$exp_week <- as.factor(beh_pairs$exp_week)

```

### Linear Mixed Model
```{r}
mixed_mod_agg <- lmer(total_aggresion ~ exp_week + (1|housing_tank_ID), data = beh_pairs)
anova(mixed_mod_agg)
```
*RESULTS*: We can see that time is a significant predictors of aggression within pairs. We are interested with pair-wise comparisons different, so we follow up with a Post-hoc Tukey analysis.

```{r}
post.hoc_agg <- glht(mixed_mod_agg, linfct = mcp(exp_week = 'Tukey'))
summary(post.hoc_agg)
```

R-squared values of the best fit model:
```{r}
r.squaredGLMM(mixed_mod_agg)
```
```{r}
emm_mix_mod_agg <- emmeans(mixed_mod_agg, ~exp_week)
pairs(emm_mix_mod_agg, simple = "each")
```

>For manuscript: GLMM results showed experimental week was significant predictors of total aggression (F-value=8.9786, NumDf=4, DenDf=94.795, p-value<0.001). Tukey post-hoc comparison showed that weeks 1, 2, and 3 were significantly different from week 5.


### Figure
```{r, include=FALSE}
beh_pairs_totaggression <- beh_pairs_totaggsub[beh_pairs_totaggsub$behaviour_type == "total_aggresion", ]
beh_pairs_totaggression
beh_pair_totalagg_summary<- ddply(beh_pairs_totaggression, c("exp_week"),summarise,
               N    = sum(!is.na(count)),
               mean = mean(count, na.rm=TRUE),
               sd   = sd(count, na.rm=TRUE),
               se   = sd / sqrt(N) )
beh_pair_totalagg_summary

colnames(beh_pair_totalagg_summary)[3] = "count"
```

```{r, fig.height=7, fig.width=7}
gg_aggression <- ggplot(beh_pairs_totaggression, aes(x=exp_week, y=count))+
        geom_jitter(size=4, alpha=0.3, width = 0.3, colour= "violetred3") +
        geom_line(group=1, data = beh_pair_totalagg_summary, size=1, colour= "maroon3")+
        geom_pointrange(aes(ymin = count-se, ymax = count+se), data =beh_pair_totalagg_summary, size=1.2, alpha=1, colour= "maroon3")+
        theme_classic() +
        theme(axis.text = element_text(size = 18), axis.title = element_text(size = 18), legend.position = c(0.9,0.9),         legend.title=element_text(size=16),  legend.text=element_text(size=16))+
        ylab("Number of aggressive encounters + SEM") +
        xlab("Experimental week")
gg_aggression

#ggsave("total_aggression.png", height = 7, width = 7, dpi = 1200)
```


---

## Submission {.tabset}
$formula: TotalSubmission$ ~ $time + (1|HousingTankID)$

### Linear Mixed Model
```{r}
mixed_mod_sub <- lmer(total_submission ~ exp_week + (1|housing_tank_ID), data = beh_pairs)
anova(mixed_mod_sub)
```


*RESULTS*: We can see that time is a significant predictors of submission within pairs. We are interested with pair-wise comparisons different, so we follow up with a Post-hoc Tukey analysis.

```{r}
post.hoc_sub <- glht(mixed_mod_sub, linfct = mcp(exp_week = 'Tukey'))
summary(post.hoc_sub)
```

>For manuscript: GLMM results showed experimental week was significant predictor of total submission (F-value=18.055 , NumDf=4, DenDf=94.791 , p-value<0.001). Tukey post-hoc comparison showed that weeks 1, 2, and 3 were significantly different from weeks 4 and 5. 

```{r}
emm_mix_mod_sub <- emmeans(mixed_mod_sub, ~exp_week)
pairs(emm_mix_mod_sub, simple = "each")
```

R-squared values of the best fit model:
```{r}
r.squaredGLMM(mixed_mod_sub)
```

### Figure
```{r, include=FALSE}
beh_pairs_totaggsub
beh_pairs_totsubmission <- beh_pairs_totaggsub[beh_pairs_totaggsub$behaviour_type == "total_submission", ]
beh_pairs_totsubmission

beh_pair_totalsub_summary<- ddply(beh_pairs_totsubmission, c("exp_week"),summarise,
               N    = sum(!is.na(count)),
               mean = mean(count, na.rm=TRUE),
               sd   = sd(count, na.rm=TRUE),
               se   = sd / sqrt(N) )
beh_pair_totalsub_summary

colnames(beh_pair_totalsub_summary)[3] = "count"

```

```{r, fig.height=7, fig.width=7}
gg_submission <- ggplot(beh_pairs_totsubmission, aes(x=exp_week, y=count))+
        geom_jitter(size=4, alpha=0.3, width = 0.3, colour= "violetred3") +
        geom_line(group=1, data = beh_pair_totalsub_summary, size=1, colour= "maroon3")+
        geom_pointrange(aes(ymin = count-se, ymax = count+se), data =beh_pair_totalsub_summary, size=1.2, alpha=1, colour= "maroon3")+
        theme_classic() +
        theme(axis.text = element_text(size = 18), axis.title = element_text(size = 18), legend.position = c(0.9,0.9),         legend.title=element_text(size=16),  legend.text=element_text(size=16))+
        ylab("Number of submissive encounters + SEM") +
        xlab("Experimental week")
gg_submission

#ggsave("total_submission.png", height = 7, width = 7, dpi = 1200)
```
## Combined Figure {.tabset}
```{r, include=FALSE}
head(beh_pairs)
head(pair_summary_totaggsub)

comb_beh_data <- gather(beh_pairs, key = "behaviour_type", value = "count", 6:7)
head(comb_beh_data)
colnames(pair_summary_totaggsub)[4] = "count"
```

```{r}
gg_comb_beh <- ggplot(comb_beh_data, aes(x=exp_week, y=count, group=behaviour_type, colour=behaviour_type))+
        geom_jitter(position = position_jitterdodge(0.3), size=4, alpha=0.3) +
        geom_line(aes(group=behaviour_type), data = pair_summary_totaggsub,  position=position_dodge(0.8), size=1)+
        geom_pointrange(aes(ymin = count-se, ymax = count+se), data =pair_summary_totaggsub, size=1.2, lwd=1.3, alpha=1,  position=position_dodge(0.8))+
        scale_color_manual(values = c( "darkorange", "grey50"), name ="behaviour", labels=c("Aggression", "Submission")) +
        theme_classic() +
        theme(axis.text = element_text(size = 18), axis.title = element_text(size = 18), legend.position = c(0.85,0.9),         legend.title=element_text(size=16),  legend.text=element_text(size=16))+
        ylab("Total aggression/submission + SEM") +
        xlab("Experimental week")

gg_comb_beh


#ggsave("../Figures/combined_behaviour.png", height = 7, width = 7, dpi = 1200)
```

