---
title: "Growth Analysis - Phenotypic Plasticity at the Initiation of Dominance hierarchies"
date: "`r Sys.Date()`"
output:
  rmdformats::downcute:
    self_contained: true
    default_style: "light"
    downcute_theme: "default"
    code_folding: hide
---

<style>
  h2{color: #DC7633 !important}
  h1{color: #5499C7 !important}
  body{background-color: white !important}
</style>

<style>
.list-group-item.active, .list-group-item.active:hover, .list-group-item.active:focus {
    color: #212F3D;
    background-color: #EDBB99;
    font-weight: bold;}
a {
    color: #5499C7;}
.nav-tabs > li.active > a, .nav-tabs > li.active > a:hover, .nav-tabs > li.active > a:focus {
    color: #DC7633;}
body {
    font-family: "Open Sans","Helvetica Neue",Helvetica,Arial,sans-serif;
    font-size: 15px;
    line-height: 1.42857143;
    color: #212F3D;}
</style>


<br/>

```{r set some defaults, echo = FALSE}

library("knitr")
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
pdf.options(useDingbats = FALSE)
options(knitr.kable.NA = '')

```


```{r load packaged, echo=FALSE}
library(ggplot2)
library(plyr)
library(dplyr)
library(multcomp)
library(tidyr)
library(lmerTest)
library(emmeans)
library(interactions)
library(kableExtra)
library(MuMIn)
```

## Introduction
This Rmarkdown document contains analysis of growth and size of fish *Amphiprion percula* which were part of a 5 week experiment. Individuals were size-matched and either housed as a pair or as a singleton. A total of 30 replicates containing 30 pairs (n=60) and 30 singletons (n=30) were used in this experiment.

Initial standard lenght (mm) was measured and recorded of all fish, prior to experimental set-up  and used to create size-matched pairs and corresponding size matched singletons. Standard lengths were measured from photos using ImageJ software. At the end of the experiment, all fish were re-measured using exact same methods. Growth of all individual was calculated as 
$FinalStanardLenghth - InitialStandardLength$.

LmerTest package was used for all general linear mixed model analysis (GLMM).


## Inspecting the dataset {.tabset}
```{r, include = FALSE}
full_data <- read.csv(file = "../Data/Vizer_StrategicGrowth_R_annotated_SNPcorrected_20250425.csv", header = T)
head(full_data)
tail(full_data)
dim(full_data)
#all looks good!

table(full_data$fish_type2) #we have PS category (replicates that lost fish - needs to be removed)

subset_data <- subset(full_data, !replicate_ID == "8" & !replicate_ID == "9" & !replicate_ID == "18")
head(subset_data)
subset_data$replicate_ID #all good
table(subset_data$fish_type2) #all good
```

### List of replicates to use
```{r}
print(subset_data$replicate_ID)
table(subset_data$fish_type2)
```
<br/>
<br/>



### Mean growth of each fish category
```{r}
group_by(subset_data, fish_type2) %>%
  summarise(
    count = n(),
    mean = mean(growth, na.rm = TRUE),
    sd = sd(growth, na.rm = TRUE))
```



---

## Linear-mixed model {.tabset}

We can think about growth over the 5 weeks
1) Does size of individual affected by fish category over time? <br/>

Hypothesis: Pairs (both P1 and P2) should grow more compared to singleton individuals. <br/>
Hypothesis: Within pairs, P1 should outgrow P2.


### Model I:
Size over time across three fish categories
$formula: size$ ~ $time*FishType2 + (1|replicateID)$

** the model reported in the manuscript **
```{r creating stacked size data, include=FALSE}
#needing columns: initial_length_avg + final_length_avg + fish_type2 + housing tank + replicate_ID 
size_data <- subset_data[, c("replicate_ID", "housing_tank_ID", "fish_type2", "initial_length_avg", "final_length_avg")]
head(size_data)
dim(size_data)  #81 rows, this needs to double to have it stacked correctly

size_data <- gather(size_data, key = "time", value = "size", 4:5)
head(size_data)
dim(size_data)  #162 rows, stacked correctly

size_data$time[size_data$time == "initial_length_avg"] <- "week 0"
size_data$time[size_data$time == "final_length_avg"] <- "week 5"
head(size_data)

#time needs to be a factor
size_data$time <- as.factor(size_data$time)
```

```{r}
mixed_mod_size <- lmer(size ~ time*fish_type2 + (1|replicate_ID), data = size_data)
anova(mixed_mod_size)
```
### Post-Hoc comparisons
```{r}
emm <- emmeans(mixed_mod_size, ~ time*fish_type2)

emms_tbl <- pairs(emm, simple = "each") %>%
  rbind() 

emms_tbl%>%
   kbl(col.names = c("Fish Category", "Time", "Contrast", "Estimate", "SE", "df", "t ratio", "p-value"), caption = "Simultaneous pair-wise post-hoc comaprisons of size of fish categories (P1, P2, S) and time (week 0, week 5).", align = "c", digits = 3) %>%
  kable_classic(full_width = F, html_font = "Cambria")%>%
  column_spec(1:3, bold = T) %>%
  row_spec(0, color = "white", background = "black",bold = T)

```

R-squared values of the best fit model:
```{r}
r.squaredGLMM(mixed_mod_size)
```

```{r}
cat_plot(mixed_mod_size, pred = time, modx=fish_type2, geom="line", interval=TRUE) 
```




```{r}
mixed_mod_size_s <- step(mixed_mod_size,reduce.fixed = FALSE, reduce.random = FALSE)
plot(mixed_mod_size_s)
```

*RESULTS:*We can see that time and fish category and the interaction between them are significant predictors of size of an individual. Given the interaction between the 2 explanatory variables are significant, we should not report main effects. 

>For manuscript: GLMM results showed that the interaction between fish category (P1, P2, vs S) and experimental week were significant predictors of size (F-value=10.534 , NumDf=2, DenDf=130, p-value<0.0001). Simultaneous pairwise comparisons using Tukey’s HSD test indicated that there was no statistical significance at week 0 among the three fish categories (Table x), meaning all fish started size matched with no significant difference. On the contrary, simultaneous pairwise comparisons using Tukey’s HSD test indicated significant differences among the three fish categories at the end of the experiment (Table x). All fish grew over the 5 week period (Table x), and P1 grew significantly more compared to P2 and S. P2 and S individuals showed no significant difference in average SL (standard length) at the end of the experiment (Table x).

Type III Analysis of Variance Table with Satterthwaite's method
                Sum Sq Mean Sq NumDF DenDF F value    Pr(>F)    
time            334.54  334.54     1   130 159.794 < 2.2e-16 ***
fish_type2      129.93   64.96     2   130  31.030 9.612e-12 ***
time:fish_type2  44.11   22.05     2   130  10.534 5.757e-05 ***

### Figures
```{r summary statistics, include=FALSE}
size_data_summary <- ddply(size_data, c("time", "fish_type2"),summarise,
               N    = sum(!is.na(size)),
               mean = mean(size, na.rm=TRUE),
               sd   = sd(size, na.rm=TRUE),
               se   = sd / sqrt(N) )
size_data_summary
```

```{r setting up data for figure, include=FALSE}
head(size_data)
size_data_summary
colnames(size_data_summary)[4] = "size"
```


```{r size figure over time, fig.height=7, fig.width=7}
gg_size <- ggplot(size_data, aes(x=time, y=size,group=fish_type2, colour=fish_type2))+
        geom_jitter(position = position_jitterdodge(0.8), size=4, alpha=0.25) +
        geom_line(aes(group=fish_type2), data = size_data_summary,  position=position_dodge(0.8), size=1)+
        geom_pointrange(aes(ymin = size-se, ymax = size+se), data =size_data_summary, size=1.2, alpha=1, position=position_dodge(0.8))+
        scale_color_manual(values = c( "magenta3","purple3", "grey50"), name ="social position") +
        theme_classic() +
        theme(axis.text = element_text(size = 18), axis.title = element_text(size = 18), legend.position = c(0.85,0.9),         legend.title=element_text(size=16),  legend.text=element_text(size=16))+
  scale_x_discrete(labels = c("week 0" = "1", "week 5" = "5")) +
        ylab("Standard length (mm) + SEM") + ylim(8,25)+
        xlab("Experimental week")
gg_size

#ggsave("size-20250428-social-rank.png", height = 7, width = 7, dpi = 1200)
```

```{r setting up growth figure, include=FALSE}
growth_summary <- ddply(subset_data, c("fish_type2"),summarise,
               N    = sum(!is.na(growth)),
               mean = mean(growth, na.rm=TRUE),
               sd   = sd(growth, na.rm=TRUE),
               se   = sd / sqrt(N) )
growth_summary
colnames(growth_summary)[3] = "growth"
```

```{r, fig.align = "center", fig.height=7, fig.width=4}
gg_growth <- ggplot(subset_data, aes(x=fish_type2, y=growth, colour=fish_type2))+
        geom_boxplot(width=0.3, fill= c("magenta3","purple3", "grey50"), alpha=0.3, position = position_dodge(0.3))+
        geom_jitter(size=4, alpha=0.3, position = position_dodge(0.3)) +
        geom_pointrange(aes(ymin = growth-se, ymax = growth+se), data =growth_summary, size=1.2, position = position_dodge(0.3), alpha=1)+
        scale_color_manual(values = c("magenta3","purple3", "grey50")) +
        theme_classic() +
        theme(axis.text = element_text(size = 24), axis.title = element_text(size = 24), legend.position = "none",         legend.title=element_text(size=24),  legend.text=element_text(size=24))+
        ylab("Growth (mm/5 weeks)") +
        xlab("social position")+
        geom_hline(yintercept=4.33, linetype="dashed", lwd=1.2, color = "magenta3")+
        geom_hline(yintercept=1.89, linetype="dashed", lwd=1.2, color = "purple3")+
        geom_hline(yintercept=2.41, linetype="dashed", lwd=1.2, color = "grey50")
gg_growth

library(cowplot)
# Combine growth, appetite, and behaviour
single <- plot_grid(gg_growth,
                           labels = c("A1"),
                           ncol = 1)


#ggsave("../Figures/growth-20250506.png", height = 7, width = 4, dpi = 1200)
```

```{r}
## combining all 3 plots
library(cowplot)
# Combine growth, appetite, and behaviour
combined_plot <- plot_grid(gg_size, gg_min_max_appetite, gg_comb_beh,  ##2 of these plots made in other 2 markdowns
                           labels = c("A", "B", "C"),
                           ncol = 1, nrow = 3, align = "v",  
                           rel_heights = c(1, 1, 1))

print(combined_plot)

#ggsave("../Figures/combined-phenotypic-fig-20250506-social-position.png", combined_plot,   height = 21, width = 6, dpi = 1200)  # Height scaled up (7 × 3)

```

